---
title: "CS544Final_Rapisardi"
author: "Giulia Rapisardi"
date: "12/17/2020"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spotify Top 50 2019

**Objective** 

To gather information on songs that are popular and see if there are any trends or common aspects of the songs that make them likeable, and therefore put them on the top 50 charts. We also will be looking at several statistics to see if any genres have similar traits.


### 1.Data Preparation

#### 1.1 Importing Data
The data set contains the top 50 most listened songs in the world by Spotify. I found
the data set on Kaggle and it can be downloaded at: 

                               (https://www.kaggle.com/leonardopena/top50spotify2019)
                      
I used this URL to read the data into R.

```{r Importing Data}
spotify <- read.csv('~/Desktop/BU/CS 544/Final Project/SpotifyTop50.csv', header = TRUE)
head(spotify)
```

The data has 14 variables, all giving details about the song and artist. Three of them are categorical and the remaining 11 are numerical. Most variables both categorical and numerical will be used while performing analysis.


#### 1.2 Pre-Processing the Data

The data set was very specific to the artists and with the variables being majority
numerical I wanted to group them into subgroups to make analysis easier.
The first column I addressed was the Genre column. There were many sub genres
that we could categorize into one genre such as converting Australian Pop to Pop and Country Rap to Rap, and so forth.

Second, I renamed the ‘X’ column as rank, to put the rows in order from most to least popular.
```{r, echo = FALSE} 
names(spotify)[names(spotify) == "X"] <- "Rank"
names(spotify)

```

This is all documented in my SCRUM document where I go through each change step by step.

*Genre*:

Contains the genre of the track. The data set has a wide variety of genres that would be difficult to run analysis on, so I grouped them into more general genres and so that there are not as many attributes in the column. The final attributes of the column are *EDM, Hip Hop, Latin, Pop, R&B, Rap, and Reggaeton*.

### 2.Analysis of Data

#### 2.1 Top 50 Genres
Music genres are what categorize artists and allow fans to explore others like them. While looking at this data and after preprocessing something that because of interest was to see what genres were more popular and if the genre made a difference in the likelihood of a song being popular. The bar graph shows the distribution of songs genres and the frequency of those songs amongst the data. The pie chart tells us the proportion of those genres along with their corresponding percentages. These statistics can help to determine what genres were more popular in 2019 and which in this case we can conclude that pop music was more popular.

 

```{r genre visuals, echo=FALSE}
genre <- spotify$Genre
table(genre)

barplot(table(genre), 
        col = 'green3', 
        las = 2,
        ylim = c(0,30), 
        xlab = 'Music Genre', 
        ylab = 'Frequencies of Genre',
        main = 'Frequency of Music Genres')

slice.labels <- names(table(genre))
slice.percent <- round(table(genre) / nrow(spotify) * 100)
slice.labels <- paste(slice.labels, slice.percent)
slice.labels <- paste(slice.labels, '%', sep = ' ')
pie(table(genre),
    labels = slice.labels,
    col = c('lightblue', 'orange', 'yellow', 'lightpink', 'honeydew3', 'plum', 'palegreen'),
    main = 'Genres Occupying the Top 50 Charts 2019')

```

#### 2.2 Tempo Influence
The tempo of a song tells us how fast and upbeat it is. This can have a big influence on how popular a song is. When initially computing the frequencies, we can see that we have a large range of BPM in this data set.

```{r bpm frequency table, echo=FALSE}
bpm <- spotify$Beats.Per.Minute
table(bpm)

```

Displayed in the bar graph below 176 was the most popular BPM from the data set, which is a very high value for this variable especially since it has a mean of 120, and maximum value of 190. But, from this individual analysis I was not able to see if there was any direct correlation between BPM and the popularity of a song as there is no prominent trend or distribution as shown below.

```{r bpm barplot, echo=FALSE}
options(scipen = 999)
barplot(table(bpm), 
        xlab = 'BPM', 
        ylab = 'Frequence of BPM',
        las = 2,
        col = 'green3', 
        xlim = c(1,37), 
        ylim = c(0,7), 
        main = 'BPM Frequencies',
        beside = TRUE)

```
To further look into the variable, the Central Limit Theorem was processed which states, that the sample means from a sample of the population will have a normal distribution. Below, determines if as the sample size increases, we get a normal distribution for this variable.

```{r bpm CLT, echo=FALSE}
bpm.mean <- mean(spotify$Beats.Per.Minute)
bpm.sd <- sd(spotify$Beats.Per.Minute)

set.seed(123)
samples <- 1000
sample.size <- 10
xbar <- numeric(samples)

par(mfrow = c(2,2))

for (size in c(90, 100, 150, 160)) {
  for (i in 1:samples) {
    xbar[i] <- mean(sample(spotify$Beats.Per.Minute, size = size,
                           replace = TRUE))
  }
  
  hist(xbar,
       xlab = 'BPM',
       ylab = 'Frequency of Song BPM',
       col = 'firebrick1',
       main = paste("Sample Size =", size))
  
  cat("Sample Size = ", size, " Mean = ", round(mean(xbar)),
      " SD = ", sd(xbar), "\n")
}

```
For this test a large sample size is used to compare what the distribution for this data would look like if it was on the scale of Billboard Top 200 or another popular chart. We can clearly see that as the sample size increases the data starts to take on the shape of a normal distribution. The mean and standard deviations for the sample sizes are as follows.


Overall, we can say that if this data set were to be larger, let’s say the top 200 songs, the BPM distribution for the data set would begin to normalize itself and create a normal distribution curve.

#### 2.3 Rank Relationships
Although all 50 songs from the dataset were great there were obviously some that were deemed better according to listeners. After analyzing the genre and tempo of the songs in the dataset, the urge to find if there were any other influential variables to the Rank seemed important to understanding the data. This sections focus was mainly on the variable Rank and the correlation it may have with other variables.

##### 2.3.1 *Rank and Danceability*
The first variable of interest is Danceability, this variable tells us how easy it is to dance to a song. The higher the value the easier it is to dance. First, I did some basic analysis using the summary function.

```{r rank danceability summary, echo=FALSE}

attach(spotify)
summary(Rank)

summary(Danceability)



```

This function gives us some basic descriptive statistics, Rank isn’t going to give us too much information and the main interest point is the summary of Danceability. Below, these statistics have been marked on the scatterplot showing the correlation between the two variables.

```{r rank and dance plot, echo=FALSE}
plot(Danceability,Rank,  main = 'Relationship Between Rank and Danceability', col = 'dodgerblue')
```
This scatterplot shows a positive correlation between the two variables, stating that as the danceability increases the rank increases or gets higher in number. It can also be infered that you are less likely to be able to dance to a song if the Rank is lower.

##### 2.3.2 *Rank and Energy*
The next variable that was used to find an influence on the rank was energy. Energy is how energetic a song is, the higher the energy the more energetic the song is. The same function was used to find basic descriptive statistics on the variable. Below is the output from the summary function.

```{r rank and energy summary, echo=FALSE}
summary(Rank)

summary(Energy)
```
As seen previously rank is the same whereas the output for energy differs from that of Danceability. The range of energy is smaller causing the 1st and 3rd quartiles to shift. These quartiles represent the median of the lower and upper half of the data. Below we can see the statistics marked on the scatterplot of the data.
```{r rank and energy plot, echo=FALSE}
plot(Energy, Rank, main = 'Relationship Between Rank and Energy', col = 'blueviolet')
```

  After searching for a pattern in the data there was no conclusion on whether or not the energy of a song influences the rank. The data in the scatterplot is distributed in no specific pattern making it hard to find a trend line.

  The question of ‘What makes a good song?’ still goes unanswered but from this analysis we can dig deeper into what other variables will influence the ranking of songs. Although, we were able to conclude that danceability certainly has an impact on the popularity of a song and may even be linked to a specific genre.
  As seen in previous analysis Pop music consumed 50% of the dataset. Since there are more songs with a danceability value larger than the mean (71.38), we could make the inference that Pop songs on the top 50 charts are more likely to have a higher-than-average danceability value than other genres. Since energy gave us no correlation, we cannot create any inferences from the data as the pattern is unknown.


#### 2.4 Song Length Correlation with Rank
Over the years, there have been studies that impose the idea that the human attention span is getting shorter. This influenced the music industry as some artists have shortened the length of their songs to keep the listener engaged and wanting to hear the chorus and hook as fast as they can. Is this portrayed in the data? By looking at the variable Length and using the Cumulative Distribution Function we can see the distribution of the songs, identifying them by their ranks and observing the relationship.

To gauge this the best, the standard deviation in reference to the mean will give the most accurate reading of the data’s distribution. After observing the data within three standard deviations of the mean, it was discovered that 68% of the data fell within one standard deviation which for this data is between 161 to 240 seconds.

Seeing that the data tends to distribute itself within one standard deviation of the data, but is this distribution normal? If we were to add more songs to the data set would this become clearer and create a prominent trend in the dataset? Using the Central Limit theorem like previously processed, we hope to understand the distribution of beats per minute in the data. The samples of the population are centered around the sample mean; in this case the sample mean is 201.

```{r song length vs rank CLT, echo=FALSE}
set.seed(123)
samples <- 1000
sample.size <- 10
xbar <- numeric(samples)

par(mfrow = c(2,2))

for (size in c(75, 100, 125, 150)) {
  for (i in 1:samples) {
    xbar[i] <- mean(sample(spotify$Length., size = size,
                                replace = TRUE) )
  }
  
  hist(xbar,
       xlab = 'Song Length',
       ylab = 'Frequency of Song Length',
       col = 'tan1',
       main = paste("Sample Size =", size))
  
  cat("Sample Size = ", size, " Mean = ", round(mean(xbar)),
      " SD = ", sd(xbar), "\n")
}

```
The above histograms show that the larger the sample size the more the distribution mimics that of a normal distribution. The larger sample sizes were chosen to relate to those of alternative music charts such as Billboard Top 200. In those charts you have a larger data set to work with which would increase the probability of a normal distribution. Below is the mean and standard deviation for the data and respective sample sizes.

From the above analysis it is clear that the distribution of songs and their length falls within a certain range. Referring back to the genres of the dataset we can make an inference about whether or not song length is influenced by genre. Since 50% of our data falls into the Pop category it would be more likely that a pop song falls within one standard deviation of the mean than any other genre. Since pop songs seemed to be more popular in 2019 there could be a relationship that pop songs tend to stay within a range of 161-240 seconds to keep their audience engaged, therefore becoming more popular. If we were to work with a larger data set with a larger variation of genres and songs, we might be able to make a conclusion on the relationship between genre and song length and if they would be normally distributed.

#### 2.5 Picking a Song from the Top 10
Now that we have analyzed the top 50 songs, this section will mainly be focused on the top 10 songs and their uniqueness using simple random sampling.

Say the names of all the songs from the data set were placed in a bag and 10 of the 50 songs were picked randomly without replacement. What are the odds that you would pick one of the songs with a rank between 1 and 10? When the analysis was performed, two of the top 10 songs were chosen as shown below.
```{r without replacement setup, include=FALSE}
library(sampling)
set.seed(1)
s1 <- srswor(10, nrow(spotify))
s1
s1[1:10]
s1[s1 != 0]

# without replacement
rows1 <- (1:nrow(spotify))[s1 != 0]
rows1

# frequency without replacement
sample.1 <- spotify[rows1, ]
sample.1  
which(sample.1$Rank <= 10)

```

```{r without replacement output, echo=FALSE}
sample.1[c(1,2), ]
```

Now, we want to perform the same analysis but with replacement. When we perform this analysis, we only get one record back.
```{r with replacement setup, include=FALSE}
### with replacement
s2 <- srswr(10, nrow(spotify))
s2
s2[1:10]
s2[s2 != 0]

# with replacement
rows2 <- (1:nrow(spotify))[s2 != 0]
rows2 

# frequency with replacement
sample.2 <- spotify[rows2, ]
sample.2  
which(sample.2$Rank <= 10)

sample.2[1, ]  

```


```{r with replacement output, echo=FALSE}
sample.2[1, ]  
```
Initially looking at these results you may think that it seems backwards. But, if you are picking the songs and replacing them, when you receive a song that is in the top 10 of the data set you put it back and have a chance of picking that song again, decreasing the chance of variability in the sampling. Whereas if you are not replacing the songs after picking them, although your likelihood of selecting a song in the top 10 decreases, the likelihood of picking a different song in the top 10 is more likely.

These results tell us that the probability of picking a song from the top 10, without replacement is 10% and the probability of picking a song from the top 10 with replacement is 20%.

#### 2.6 Artist Info
```{r stringr setup, include=FALSE}
library(stringr)
```

Prior analysis has been focused on the Rank of a song, now we will pull some basic information on the Artists of the data set. With a variation of artist names, being able to easily sort and locate artists of a certain name or starting with a certain letter can be helpful when determining information on that artist and their song.

Starting off simple, *how many artists names start with an A?*
```{r start A, echo=FALSE}
table(str_subset(spotify$Artist.Name, "^A"))
```

From the 50 songs in the dataset only three artist names start with an ‘A’ and one of those names occurs twice. The probability of randomly selecting an artist with a name starting with ‘A’ is 8%.

What about artists that have names that *end with an A*?
```{r end with A, echo=FALSE}
table(str_subset(spotify$Artist.Name, "A$"))
```

From the 50 songs in the data set only three artists have names that end with an ‘A’ and there were no duplicates in this table. The probability of randomly selecting an artist with a name ending in ‘A’ is 6%.

This function was also performed to find artists with *names starting or ending with ‘R’*
```{r start and end with R, echo=FALSE}
### How many artists names start with R?
table(str_subset(spotify$Artist.Name, "^R"))
### How many artists names end with R?
table(str_subset(spotify$Artist.Name, "R$"))
```
When performed we can see that there was only one artist that’s name started with the letter ‘R’ and there were no artists that’s name ended with ‘R’.

This function allowed for good exploration of the data and helped to get a better understanding of the variables. In a larger data set it could be useful for grouping data and running analysis on those groups specifically.

#### 2.7 Top 10 Features
Reverting back to the top 10 songs of the dataset we now understand the likelihood of a song randomly being chosen from the top 10, but, are there certain aspects that these songs have in common that make them popular?

Using the tidyverse package, the first 10 rows were selected and reassigned to the tibble spotify.top.10 which will be used for the remainder of the analysis in this section.

To compute all of the descriptive statistics at once we use the summary function, which produced the output below.


```{r tibble setup, include=FALSE}
library(tidyverse)

# identify the top 10 tracks
spotify.top.10 <- spotify[1:10, ]

# convert the top 10 tracks to a tibble
as_tibble(spotify.top.10)
```
```{r top 10 summary, echo=FALSE}
spotify.top.10 %>% summary
```

Paying attention to the variables used for previous analysis we see that the average BPM increased from 120 to 132, the average length of a song decreased from 201 seconds to 196 seconds, and the average danceability of a song decreased from 71.38 to 67.80.

Seeing that the BPM increased for the subset of data, further analysis will aid in locating specific songs correlating to their BPM. First, *which songs from the data have a BPM larger than 100?*

```{r bpm > 100, echo=FALSE}
# BPM greater than 100? 
spotify.top.10 %>%
  select(Track.Name, Rank, Artist.Name, Genre, Beats.Per.Minute) %>%
  filter(Beats.Per.Minute > 100)
```
Only 9 of the 10 songs from the data have a BPM > 100.

*Which songs have a BPM larger than 125?*
```{r bpm > 125, echo=FALSE}
# BPM greater than 125?
spotify.top.10 %>%
  select(Track.Name, Rank, Artist.Name, Genre, Beats.Per.Minute) %>%
  filter(Beats.Per.Minute > 125)
```
Five of the songs from the data have a BPM > 125. 

*Which songs have a BPM larger than 150?*
```{r bpm > 150, echo=FALSE}
# BPM greater than 150?
spotify.top.10 %>%
  select(Track.Name, Rank, Artist.Name, Genre, Beats.Per.Minute) %>%
  filter(Beats.Per.Minute > 150)
```
Two of the songs from the data have a BPM > 150.


From this we can see that on average half of the data from this subset has a higher BPM than the average of the whole population of the Spotify dataset, therefore the increase in the average BPM is due to the increase in songs with higher BPM. From this we can infer that songs that are more popular tend to have a higher BPM.

Now we can address the length of the top 10 songs. Using similar functions, we will determine which songs fall within certain time intervals.

*Which songs are longer than 3 minutes (180 seconds)?*
```{r length > 180, echo=FALSE}
# we are interested to see if the top 10 songs are longer in length (or longer than 3 minutes, 180 seconds)
spotify.top.10 %>%
  select(Track.Name, Rank, Artist.Name, Genre, Length.) %>%
  filter(Length. > 180)
```
Seven of the songs from the data have a length > 180 seconds.

*What songs are longer than 3 1⁄2 minutes (210 seconds)?*
```{r lenght > 210, echo=FALSE}
# longer than 3 1/2 minutes (210 seconds)?
spotify.top.10 %>%
  select(Track.Name, Rank, Artist.Name, Genre, Length.) %>%
  filter(Length. > 210)
```
Two of the songs from the data have a length > 210 seconds. 

*What songs are longer than 4 minutes (240 seconds)?*
```{r length > 240, echo=FALSE}
# longer than 4 minutes (240 seconds)?
spotify.top.10 %>%
  select(Track.Name, Rank, Artist.Name, Genre, Length.) %>%
  filter(Length. > 240)
```
Only one song from the data has a length > 240 seconds.

Overall, the decrease in song length for the subset of data is influenced by the overall decrease in song lengths. From this we can infer that shorter songs are more likely to be popular.

Assuming that our inference is correct, that shorter songs are more likely to be popular, does the length influence the genre of a song? Seeing as the data had more Pop songs that were popular further analysis will conclude what the average song length each genre holds.

```{r avg length/genre, echo=FALSE}
# Regroup the data by genre, and find the average length of songs for that genre
avg_length <- spotify.top.10 %>%
  select(Track.Name, Rank, Artist.Name, Genre, Length.) %>%
  group_by(Genre) %>%
  summarise( avg_length = mean(Length.) )

avg_length
```
Because we are working with a subset of data not all genres are going to be included. The results contain four of the seven genres of which Reggaeton songs had the longest average length, Pop following as the second longest average length, then Rap, and Hip Hop had the shortest average length. These values tell us that for these genres a song that averages this length is more likely to be popular and fall in the top 10 charts.

Since there is an average length that a genre might hold that makes it popular, is there an average BPM that a genre has that would make it popular?

```{r avg bpm / genre, echo=FALSE}
# Regroup the data by genre, and find the average BPM for songs for that genre
avg_bpm <- spotify.top.10 %>%
  select(Track.Name, Rank, Artist.Name, Genre, Length.) %>%
  group_by(Genre) %>%
  summarise( avg_bpm = mean(Beats.Per.Minute) )

avg_bpm
```

The data here shows that all of the genres in the subset have the same average BPM. Therefore, we can conclude that the BPM for these genres does not influence the popularity or rank of a song.

#### 4. Conclusion
There were a lot of interesting statistics and trends found in this dataset. I chose this dataset specifically because I have always had a passion and interest in music and a career goal of mine would be to do analysis for Spotify and find trends in data that would help artists be successful. So, taking a look at some past record-breaking artists lets discuss the results.
Pop music was one of the most popular genres in 2019, with an average BPM of 120 and song length of 198 seconds. This combined with the inference between rank and danceability a final analysis would be that Pop songs, or songs that follow the outline of a pop song are more likely to have a higher rank and be popular.
This analysis was derived from a smaller data set of only 50 rows so a larger dataset would allow for further analysis and information about songs and what makes a song a hit.



















